Methods for Large Scale Learning to Rank

### 2014-09-25 report/Chapters/BenchmarkResults.tex
I -> we

The Yahoo! Learning to Rank Challenge consists of two tracks that uses the two data sets respectively: a standard Learning to Rank track and a transfer learning track where the goal was to learn a specialized ranking function that can be used for a small country by leveraging a larger training set of another country. For this experiment, I will only look at the standard Learning to Rank data set, because transfer learning is a separate research area that is not included in this thesis.\\
\begin{table}[!h]
\begin{tabular}{l|lll}
 & Train & Validation & Test \\ 
 \hline
\# of queries & 19,994 & 2,994 & 6,983 \\ 
\# of documents & 473,134 & 71,083 & 165,660 \\ 
\# of features & 519 & 519 & 519 \\ 
\end{tabular}
\caption{Yahoo! Learning to Rank Challenge data set characteristics, as described in the overview paper \cite{Chapelle2011a}}
\label{tab:yahoo_characteristics}
\end{table}\\
Both \ac{NDCG} and \ac{ERR} are measured as performance metrics, but the final standings of the challenge were based on the \ac{ERR} values. Model validation on the Learning to Rank methods participating in the challenge is performed using a train/validation/test-set split following the characteristics shown in Table \ref{tab:yahoo_characteristics}. Competitors could train on the training set and get immediate feedback on their performance on the validation set. The test set performance is used to create the final standings and is only measured after the competition has ended to avoid overfitting on the test set. The large number of documents, queries and features compared to other benchmark data sets makes the Yahoo! Learning to Rank Challenge data set interesting. Yahoo did not provide detailed feature descriptions to prevent competitors to get detailed insight in the characteristics of the Yahoo data collection and features used at Yahoo. Instead high level descriptions of feature categories are provided. The following categories of features are described in the challenge overview paper \cite{Chapelle2011a}:\\
\begin{description}
\item[Web graph]{Quality and popularity metrics of web documents, e.g. PageRank \cite{Page1999}}.
\item[Document statistics]{Basic document statistics such as the number of words and url characteristics.}
\item[Document classifier]{Results of various classifiers on the documents. These classifiers amongst others include: spam, adult, language, main topic, and quality classifiers.}
\item[Query]{Basic query statistics, such as the number of terms, query frequency, and click-through rate.}
\item[Text match]{Textual similarity metrics between query and document. Includes \ac{TF-IDF}, BM25 \cite{Robertson2009} and other metrics for different sections of the document.}
\item[Topical matching]{These features go beyond similarity at word level and compute similarity on topic level. For example by classifying both the document and the query in a large topical taxonomy.}
\item[Click]{Click-based user feedback.}
\item[External references]{Document meta-information such as Delicious\footnote{https://delicious.com/} tags}
\item[Time]{Document age and historical in- and outlink data that might help for time sensitive queries.}
\end{description}

@@

I will describe the performance of the algorithms that were evaluated by the LETOR team on LETOR 3.0 after publication of the LETOR paper by Qin et al. \cite{Qin2010}, as listed on the LETOR website\footnotemark[2] by comparing their performance with the ListNet baseline. We will consider those methods to be better than ListNet when they beat the ListNet baseline in at least four of the seven data sets in \ac{NDCG} value. Note that this does not necessarily imply that these methods would have scored a higher \ac{NDCG} winning number than ListNet. Table \ref{tbl:LETOR_ListNet} shows the ListNet performance on the seven LETOR data sets in terms of \ac{NDCG}@10 and \ac{MAP}.
\begin{table}[!h]
\begin{tabular}{l|ll}
 & \ac{NDCG}@10 & \ac{MAP} \\ 
 \hline
TD2003 & 0.348 & 0.2753 \\ 
TD2004 & 0.317 & 0.2231 \\ 
NP2003 & 0.801 & 0.6895 \\ 
NP2004 & 0.812 & 0.6720 \\ 
HP2003 & 0.837 & 0.7659 \\ 
HP2004 & 0.784 & 0.6899 \\ 
OHSUMED & 0.441 & 0.4457 \\ 
\end{tabular}
\caption{Performance of ListNet on LETOR 3.0}
\label{tbl:LETOR_ListNet}
\end{table}

>>>>

The Yahoo! Learning to Rank Challenge consists of two tracks that uses the two data sets respectively: a standard Learning to Rank track and a transfer learning track where the goal was to learn a specialized ranking function that can be used for a small country by leveraging a larger training set of another country. For this experiment, we will only look at the standard Learning to Rank data set, because transfer learning is a separate research area that is not included in this thesis.\\
\begin{table}[!h]
\begin{tabular}{l|lll}
 & Train & Validation & Test \\ 
 \hline
\# of queries & 19,994 & 2,994 & 6,983 \\ 
\# of documents & 473,134 & 71,083 & 165,660 \\ 
\# of features & 519 & 519 & 519 \\ 
\end{tabular}
\caption{Yahoo! Learning to Rank Challenge data set characteristics, as described in the overview paper \cite{Chapelle2011a}}
\label{tab:yahoo_characteristics}
\end{table}\\
Both \ac{NDCG} and \ac{ERR} are measured as performance metrics, but the final standings of the challenge were based on the \ac{ERR} values. Model validation on the Learning to Rank methods participating in the challenge is performed using a train/validation/test-set split following the characteristics shown in Table \ref{tab:yahoo_characteristics}. Competitors could train on the training set and get immediate feedback on their performance on the validation set. The test set performance is used to create the final standings and is only measured after the competition has ended to avoid overfitting on the test set. The large number of documents, queries and features compared to other benchmark data sets makes the Yahoo! Learning to Rank Challenge data set interesting. Yahoo did not provide detailed feature descriptions to prevent competitors to get detailed insight in the characteristics of the Yahoo data collection and features used at Yahoo. Instead high level descriptions of feature categories are provided. The following categories of features are described in the challenge overview paper \cite{Chapelle2011a}:\\
\begin{description}
\item[Web graph]{Quality and popularity metrics of web documents, e.g. PageRank \cite{Page1999}}.
\item[Document statistics]{Basic document statistics such as the number of words and url characteristics.}
\item[Document classifier]{Results of various classifiers on the documents. These classifiers amongst others include: spam, adult, language, main topic, and quality classifiers.}
\item[Query]{Basic query statistics, such as the number of terms, query frequency, and click-through rate.}
\item[Text match]{Textual similarity metrics between query and document. Includes \ac{TF-IDF}, BM25 \cite{Robertson2009} and other metrics for different sections of the document.}
\item[Topical matching]{These features go beyond similarity at word level and compute similarity on topic level. For example by classifying both the document and the query in a large topical taxonomy.}
\item[Click]{Click-based user feedback.}
\item[External references]{Document meta-information such as Delicious\footnote{https://delicious.com/} tags}
\item[Time]{Document age and historical in- and outlink data that might help for time sensitive queries.}
\end{description}

@@

We will describe the performance of the algorithms that were evaluated by the LETOR team on LETOR 3.0 after publication of the LETOR paper by Qin et al. \cite{Qin2010}, as listed on the LETOR website\footnotemark[2] by comparing their performance with the ListNet baseline. We will consider those methods to be better than ListNet when they beat the ListNet baseline in at least four of the seven data sets in \ac{NDCG} value. Note that this does not necessarily imply that these methods would have scored a higher \ac{NDCG} winning number than ListNet. Table \ref{tbl:LETOR_ListNet} shows the ListNet performance on the seven LETOR data sets in terms of \ac{NDCG}@10 and \ac{MAP}.
\begin{table}[!h]
\begin{tabular}{l|ll}
 & \ac{NDCG}@10 & \ac{MAP} \\ 
 \hline
TD2003 & 0.348 & 0.2753 \\ 
TD2004 & 0.317 & 0.2231 \\ 
NP2003 & 0.801 & 0.6895 \\ 
NP2004 & 0.812 & 0.6720 \\ 
HP2003 & 0.837 & 0.7659 \\ 
HP2004 & 0.784 & 0.6899 \\ 
OHSUMED & 0.441 & 0.4457 \\ 
\end{tabular}
\caption{Performance of ListNet on LETOR 3.0}
\label{tbl:LETOR_ListNet}
\end{table}

### 2014-09-25 report/Chapters/CrossBenchmarkComparison.tex
I -> we

Several studies make a small start in considering Learning to Rank methods performance over multiple benchmark data sets. Gomes et al. \cite{Gomes2013} analysed ranking accuracy of a set of models on both LETOR 3.0 and LETOR 4.0. Busa-Fekete et al. \cite{Busa-Fekete2013} compared the accuracy of a small set of models over the LETOR 4.0 data sets, both MSLR data sets, both Yahoo! Learning to Rank Challenge data sets and the OHSUMED dataset from LETOR 3.0. To my knowledge, no structured meta-analysis on ranking accuracy has been conducted where evaluation results on several benchmark collections are taken into account. With a meta-analysis we will compare the performance of Learning to Rank methods across the Learning to Rank benchmark data sets described in foregoing sections.

@@

\section{Conclusions}
We will now look back on our first research question using the knowledge gained with the cross-benchmark comparison.
\begin{description}
\item[RQ1] What are the best performing Learning to Rank algorithms in terms of accuracy on relevant benchmark data sets?\\
\end{description}
Although no closing arguments can be formulated on which Learning to Rank methods are most accurate, a lot of insight has been gained with the cross-benchmark comparison on which methods tend to perform better than others.\\

>>>>

Several studies make a small start in considering Learning to Rank methods performance over multiple benchmark data sets. Gomes et al. \cite{Gomes2013} analysed ranking accuracy of a set of models on both LETOR 3.0 and LETOR 4.0. Busa-Fekete et al. \cite{Busa-Fekete2013} compared the accuracy of a small set of models over the LETOR 4.0 data sets, both MSLR data sets, both Yahoo! Learning to Rank Challenge data sets and the OHSUMED dataset from LETOR 3.0. To our knowledge, no structured meta-analysis on ranking accuracy has been conducted where evaluation results on several benchmark collections are taken into account. With a meta-analysis we will compare the performance of Learning to Rank methods across the Learning to Rank benchmark data sets described in foregoing sections.

@@

\section{Conclusions}
We will now look back on our first research question using the knowledge gained with the cross-benchmark comparison.
\begin{description}
\item[RQ1] What are the best performing Learning to Rank algorithms in terms of ranking accuracy on relevant benchmark data sets?\\
\end{description}
Although no closing arguments can be formulated on which Learning to Rank methods are most accurate, a lot of insight has been gained with the cross-benchmark comparison on which methods tend to perform better than others.\\

### 2014-09-25 report/Chapters/LTRMethods.tex
I -> we

\section{SmoothRank}
SmoothRank \cite{Chapelle2010} is a listwise ranking method that, in contrast to ListNet, directly optimises an information retrieval evaluation measure. SmoothRank enables direct optimisation of the non-convex and discontinuous evaluation measure by smoothing, that is, approximating the rank position. In this section, I apply the SmoothRank approximation method to the \ac{NDCG} evaluation measure for illustration, but the same procedure can be applied to \ac{MAP} or any other information retrieval measure. The smoothing function used in SmoothRank is based on the softmax activation function \cite{Bridle1990}, which is often used in neural networks. The softmax function is formally defined as shown in Equation \ref{eq:softmax_algorithm}.
\begin{equation}
p_i = \frac{e^{f_i/\sigma}}{\sum\nolimits_{j=1}^{m}e^{f_j/\sigma}}
\label{eq:softmax_algorithm}
\end{equation}

>>>>

\section{SmoothRank}
SmoothRank \cite{Chapelle2010} is a listwise ranking method that, in contrast to ListNet, directly optimises an information retrieval evaluation measure. SmoothRank enables direct optimisation of the non-convex and discontinuous evaluation measure by smoothing, that is, approximating the rank position. In this section, we apply the SmoothRank approximation method to the \ac{NDCG} evaluation measure for illustration, but the same procedure can be applied to \ac{MAP} or any other information retrieval measure. The smoothing function used in SmoothRank is based on the softmax activation function \cite{Bridle1990}, which is often used in neural networks. The softmax function is formally defined as shown in Equation \ref{eq:softmax_algorithm}.
\begin{equation}
p_i = \frac{e^{f_i/\sigma}}{\sum\nolimits_{j=1}^{m}e^{f_j/\sigma}}
\label{eq:softmax_algorithm}
\end{equation}

### 2014-09-25 report/Chapters/RelatedWork.tex
I -> we

An abstract-based manual filtering step is applied where those results are filtered that use the terms \emph{parallel} or \emph{distributed} in context to \emph{learning to rank}, \emph{learning-to-rank} or \emph{machine learned ranking}. As a last step I will filter out studies based on the whole document that only focus on efficient query evaluation and not on parallel or distributed learning of ranking functions, as those studies are likely to meet listed search terms.\\

>>>>

An abstract-based manual filtering step is applied where those results are filtered that use the terms \emph{parallel} or \emph{distributed} in context to \emph{learning to rank}, \emph{learning-to-rank} or \emph{machine learned ranking}. As a last step we will filter out studies based on the whole document that only focus on efficient query evaluation and not on parallel or distributed learning of ranking functions, as those studies are likely to meet listed search terms.\\
